Notes on agt

### 2024-08-13 agt/survey.tex
GN statement

\renewcommand{\div}{\del_{\Omega}}
\renewcommand{\PV}{\mathrm{PV}}
\renewcommand{\op}{\operatorname}

@@

Turning on this deformation leads to the deformed perturbative phase space
\beqn
\begin{tikzcd}
\ul{-2} & \ul{-1} & \ul{0} \\
& \Omega^{0,\bu}(S, \lie{g}) \ar[dd,dotted,"\omega \wedge \del"] & \\
\Omega^{2,\bu}(S,\lie{g} \otimes L^{-1}) & & \Omega^{0,\bu}(S, \lie{g} \otimes L) \\
& \Omega^{2,\bu}(S, \lie{g}) & .
\end{tikzcd}
\eeqn
A more invariant way to construct this perturbative phase space is the following.
By transgression, we can view the invariant form as a one-form on the moduli stack $\op{Bun}_G(S)$ of cohomological degree $-1$.
We can pull this one-form back to get a one-form on $S \times \op{Bun}_G(S)$ also of cohomological degree $-1$.
Integration of this one-form against $\omega$ along the fibers of $S \times \op{Bun}_G(S) \to \op{Bun}_G(S)$ yields a two-form on $\op{Bun}_G(S)$ (of cohomological degree zero).
We can, in turn, further restrict this to a two-form on $\cM_{G}(S,L)$ that we will abusively still denote by $\omega$.
The phase space is then the twisted cotangent bundle
\beqn
\T_{\omega}^* \left(\cM_{G}(S,L) \right) .
\eeqn

@@

Suppose that $V$ is a graded vector space equipped with a graded symmetric pairing of total degree zero.
Let $\{e_i\}$ be a basis for $V$.
There is a vertex algebra $\cC\cB(V)$ generated by fields $e_i(z)$ of spin one and with OPEs
\beqn
e_i(z) e_j(w) \simeq \frac{(e_i,e_j)}{(z-w)^2} .
\eeqn
This is a graded vertex algebra with grading inherited from that of $V$.
Notice when $V = \C$ with the obvious pairing this is simply the chiral boson vertex algebra.

Similarly, suppose that $U$ is a $\Z \times \Z/2$ vector space equipped with a graded totally anti-symmetric pairing $\<-,-\>$ of total degree zero.
Then we can consider the symplectic boson vertex algebra valued in $U$ that we denote $\cS\cB(U)$.
If $\{f_i\}$ is a basis for $U$ then the vertex algebra is generated by free fields $f_i(z)$ of spin $1/2$ and with OPEs
\beqn
f_i(z) f_j(w) \simeq \frac{\<f_i,f_j\>}{z-w} .
\eeqn
This vertex algebra inherits a $\Z \times \Z/2$ grading from $U$.

\begin{prop}
The vertex algebra $\VV_{S,W}$ is quasi-isomorphic to the vertex algebra
\beqn
\cC\cB(H^\bu(S,\Omega^1_S)) \otimes \cS\cB(\Pi H^\bu(S,W)[1]) \otimes T
\eeqn 
where $T$ is the vector space
\beqn
\Sym(H^\bu(S,K_S)[1])
\eeqn
equipped with the trivial vertex algebra structure.
\end{prop}

>>>>

\renewcommand{\div}{\del_{\Omega}}
\renewcommand{\PV}{\mathrm{PV}}
\renewcommand{\op}{\operatorname}
\newcommand{\CB}{\mathbb{C}\mathbb{B}}
\newcommand{\SB}{\mathbb{S}\mathbb{B}}

@@

Turning on this deformation leads to the deformed perturbative phase space
\beqn
\begin{tikzcd}
\ul{-2} & \ul{-1} & \ul{0} \\
& \Omega^{0,\bu}(S, \lie{g}) \ar[dd,dotted,"\omega \wedge \del"] & \\
\Omega^{2,\bu}(S,\lie{g} \otimes L^{-1}) & & \Omega^{0,\bu}(S, \lie{g} \otimes L) \\
& \Omega^{2,\bu}(S, \lie{g}) & .
\end{tikzcd}
\eeqn
A more invariant way to construct this perturbative phase space is the following.
The non-degenerate invariant pairing determines a four-form $\kappa$ on $BG$.
This pulls back, along the evaluation map
\beqn
\op{ev} \colon S \times \op{Bun}_G(S) \to BG 
\eeqn
to a four-form on $S \times \op{Bun}_G(S)$.
Integration of the six-form $\omega \wedge \op{ev}^*\kappa$ along $S$ yields a two-form $\omega_{\kappa} = \int_S \omega \wedge \op{ev}^*\kappa$ on $\op{Bun}_G(S)$.
We can, in turn, further restrict this to a two-form on $\cM_{G}(S,L)$ that we will abusively still denote by $\omega_{\kappa}$.
The phase space is then the twisted cotangent bundle
\beqn
\T_{\omega_{\kappa}}^* \left(\cM_{G}(S,L) \right) .
\eeqn

@@

Suppose that $V$ is a graded vector space equipped with a graded symmetric pairing of total degree zero.
Let $\{e_i\}$ be a basis for $V$.
There is a vertex algebra $\CB[V]$ generated by fields $e_i(z)$ of spin one and with OPEs
\beqn
e_i(z) e_j(w) \simeq \frac{(e_i,e_j)}{(z-w)^2} .
\eeqn
This is a graded vertex algebra with grading inherited from that of $V$.
Notice when $V = \C$ with the obvious pairing this is simply the chiral boson vertex algebra.

Similarly, suppose that $U$ is a $\Z \times \Z/2$ vector space equipped with a graded totally anti-symmetric pairing $\<-,-\>$ of total degree zero.
Then we can consider the symplectic boson vertex algebra valued in $U$ that we denote $\SB[U]$.
If $\{f_i\}$ is a basis for $U$ then the vertex algebra is generated by free fields $f_i(z)$ of spin $1/2$ and with OPEs
\beqn
f_i(z) f_j(w) \simeq \frac{\<f_i,f_j\>}{z-w} .
\eeqn
This vertex algebra inherits a $\Z \times \Z/2$ grading from $U$.

\begin{prop}
The vertex algebra $\VV_{S,W}$ is quasi-isomorphic to the vertex algebra
\beqn
\CB[H^\bu(S,\Omega^1_S)] \otimes \SB[\Pi H^\bu(S,W)[1]] \otimes T
\eeqn 
where $T$ is the vector space
\beqn
\Sym(H^\bu(S,K_S)[1])
\eeqn
equipped with the trivial vertex algebra structure.
\end{prop}

We now focus on a different, more familiar, choice of bundle data.
Suppose $S$ is a projective surface as before.
This time, we choose 
\beqn
W = \cO_S \boxtimes K_C \oplus K_S \boxtimes \cO_C .
\eeqn
and further regrade using the fact that $SU(2)$ has been broken to $U(1)$.
This six-dimensional theory is of the Dolbeault complex of the $\Z$-graded complex of holomorphic vector bundles
\beqn
\begin{tikzcd}
\ul{-2} & \ul{-1} & \ul{0} \\
& \Omega^2_{S \times C} \ar[r,"\del"] & \Omega^3_{S \times C} \\
\cO_{S} \boxtimes \Omega^1_C & & \Omega^2_S \boxtimes \cO_C .
\end{tikzcd} 
\eeqn
The BV quantization of this theory yields a factorization algebra on $S \times C$ that we will denote $\Obs_{S}^{NG}$.
Its compactification, or factorization homology, along $S$
\beqn
\int_S \Obs^{NG}_{S}
\eeqn
is a holomorphic factorization algebra on $C$.
On $C = \C$ this determines a vertex algebra that we denote $\VV_{S}^{NG}$.

\begin{prop}
The vertex algebra $\VV_{S}^{NG}$ is quasi-isomorphic to
\beqn
\CB[H^\bu(S,\Omega^1_S)] \otimes \beta\gamma[H^\bu(S,K)] \otimes T
\eeqn 
where $T$ is the vector space
\beqn
\Sym(H^\bu(S,K_S)[1])
\eeqn
equipped with the trivial vertex algebra structure.
The character of this vertex algebra (whose fugacities we explain in the proof) recovers G\"ottsche's formula for the Hodge-Poincar\'e polynomial of the Hilbert schemes of $S$:
\beqn
\chi(\VV_S^{NG}) = \prod_{m \geq 1} \prod_{k,\ell =0}^2 \left(1 - (-1)^{k+\ell} x^{k+m-1} y^{\ell + m - 1} q^m \right)^{(-1)^{k+\ell+1} h^{k,\ell}(S)} .
\eeqn
\end{prop}
\begin{proof}
Let $\til q$ be the fugacity associated to the conformal weight operator $L_0$.
Let $u$ be the fugacity for cohomological degree.
Let $v$ be the `flavor' fugacity, where $\gamma$ is of weight one and $\beta$ is opposite.
Finally, let $w$ be a fugacity corresponding to the rather trivial symmetry which scales the generators of the commutative part of the vertex algebra with weight one.

Then the single particle characters of each of the fields is computed directly as follows (here $h^{k,\ell} = h^{k,\ell}(S)$).
\begin{itemize}
\item The $\gamma$ field in the $\beta\gamma$ system valued contributes
\beqn
\frac{v}{1-\til q} \left(h^{2,0} - u h^{2,1} + u^2 h^{2,2}\right) .
\eeqn
\item The $\beta$ field in the $\beta\gamma$ system valued contributes
\beqn
\frac{v^{-1} \til q}{1- \til q} \left(u^{-2} h^{0,0} - u^{-1} h^{0,1} + h^{0,2}\right) .
\eeqn
\item The chiral boson field contributes
\beqn
\frac{\til q}{1-\til q} \left(-u^{-1} h^{1,0} + h^{1,1} - u h^{1,2} \right) .
\eeqn
\item The commutative part of the vertex algebra contributes
\beqn
w \left(-u^{-1} h^{2,0} + h^{2,1} - u h^{2,2}\right) .
\eeqn
\end{itemize}
To recover G\"ottsche's formula we specialize
\beqn
\til q = xy q, \quad u = y, \quad v = x y^{-1}, \quad w = x .
\eeqn
\end{proof}
